### 2014-09-12 manuscript.Rmd
writing

---
layout: preprint
title: "An introduction to Docker for reproducible research, with examples from the R environment"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
abstract: |
  Docker is a promising solution to

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output: 
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

<!--

Suggested paper topics include, but are not limited to:

  * novel ways to manage systems research artifacts, dependencies, workflows,
    and data collection
  * testbeds for repeatable research and/or publishing executable artifacts
  * "rich publications" that connect papers to software and data
  * repeatability in distributed and nondeterministic systems
  * mitigating obstacles to artifact sharing, repeatability, and reproduction
  * repeatability challenges specific to systems
  * measurement bias in systems experiments
  * experiences and outcomes from systems repetition and reproduction studies
  * improving scientific practice in systems research
  * incentives for producing and sharing high-quality research artifacts

Papers are expected to present significant results, insights, and/or directions
for future work, and they are expected to include substantial material that has
not already been published.

SUBMISSION INSTRUCTIONS

Submissions must be no longer than ten (10) pages; shorter papers are allowed
and encouraged.  Submissions must be written in English and follow the standard
formatting guidelines for papers appearing in Operating Systems Review (see
<http://www.sigops.org/osr.html>).  Papers must be submitted in PDF format via
the submission web site.  Submitted papers will be reviewed by the guest editor
and the review committee.  Papers will be evaluated on technical quality,
originality, relevance, and presentation.  Accepted papers will be published in
the January 2015 issue of Operating Systems Review.

-->

Reproducible research has recieved an increasing level of attention
throughout the scientific community (e.g. [Peng 2011], [Nature editorial])
and the public at large (e.g. [The Economist 2014]). All steps of the
scientific process, from data collection and processing, to analyses,
visualizations and conclusions depend ever more on computation and
algorithms, _computational reproducibility_ has recieved particular
attention [Merali 2010].  Though in principle this algorithmic dependence
should make such research easier to reproduce -- computer codes being
both more portable and potentially more precise to exchange and run
than experimental methods -- in practice this has led to an ever larger
and more complex black box that stands between what was actually done
and what is described in the literature.  Crucial scientific processes
such as replicating the results, extending the approach or testing the
conclusions in other contexts, or even merely installing the software
used by the original researchers can become immensely time-consuming if
not impossible.

# A cultural problem

It is worth observing from the outset that the primary barrier to
computational reproducibility in many domain sciences has nothing
to do with the technological approaches discussed here, but stems rather from a
reluctance to publish the code used to generating the results in the
first place ([Barnes 2010]). Despite extensive evidence to the contrary
(e.g. [Ince 2012]), many researchers and journals continue to assume that
summary descriptions or pseudo-code provide a sufficient description of
computational methods used in data gathering, processing, simulation,
visualization, or analysis.  Until such code is available in the first
place, we cannot even begin to encounter the problems that the approaches
discussed here set out to solve. As a result, few domain researchers
may be fully aware of the challenges involved in effectively re-using
published code.

A lack of requirements or incentives no doubt plays a crucial role in
discouraging sharing ([Barnes 2010], [Stodden 2013]). Nevertheless,
it is easy to underestimate the siginficant barriers raised by a lack
of familiar, intuitive, and widely adopted tools for addressing the
challenges of computational reproducibility. Surveys and case studies find
that a lack of time, more than innate opposition to sharing, discourages
researchers from providing code ([Stodden 2010], [FitzJohn 2014]).

# The four technical challenges of computational reproducibility

By restricting ourselves to studies of where code has been made
available, we may sidestep for the moment the cultural challenges
to reproducibility so that we may focus on the technical ones; in
particular, those challenges for which improved tools and techniques
rather than merely norms of behavior can contribute substantially to
improved reproducibility.

Studies focusing on code that has been made available with scientific
publications regularly find the same common issues that pose substantial
barriers to reproducing the original results or building on that code
(e.g. [Lapp 2014], [Garijo 2013], [Gilbert 2012], [Collberg 2014],
[Brown 2014]), which I attempt to summarize as follows.

<!-- Is this a permitted term? -->

1. **"Dependency Hell"**

A recent study by researchers at the University of Arizona found that
less than 50% of software could even be successfully built or installed
[Collberg 2014], and in an ongoing effort by other researchers to
replicate that study [Brown 2014]. Installing or building software
necessary to run the code in question assumes the ability to recreate
the computational environment of the original researchers.

<!-- FIXME more stuff on what this means-->

<!-- Time savings. Personal reproducibility. Collaboration. Teaching. -->

<!-- Figure 1: technical reproducibility stack
1. Operating system
2. Workflow systems 
3. Package managers 
4. Dynamic documents

-->

<!-- Omit this? -->
Differences in numerical evaluation, such as arise in floating point
arithmetic or computers or even ambiguities in standardized programming
languages ("order-of-evaluation" problems) can be responsible for
differing results between or even within the same computational platform
([Ince 2012]). Such issues make it difficult to restrict the true dependencies
of the code to higher level environments such as that of a given scripting
language, indpendent of the underlying OS or even hardware itself.

2.  **Imprecise documentation**

Documentation on how to install and run code associated with published
research is another frequent barrier to replication.  [Lapp 2014] found
this impairs a researcher's ability to install and build the software
necessary, as even small holes in the documentation were found to be
major barriers, particularly for "novices" (Sensu [Garijo 2013] --
where novices may be experts in nearby langauages but unfamiliar with
the package managers and other tools of the language involved). This
same problem is discussed in [Clark 2014].  Imprecise documentation
goes well beyond issues of the software environment itself: incomplete
documentation of parameters involved meant as few as 30% of analyses
(n=34) using the popular software STRUCTURE could be reporoduced in the
study of [Gilbert 2012].

3.  **Code rot**

Software dependencies are not static elements, but recieve regular
updates that may fix bugs, add new features or deprecate old features
(or even entire dependencies themselves).  Any of these changes can
potentially change the results generated by the code.  As some of these
changes may indeed resolve valid bugs or earlier problems with underlying
code, it will often be insufficient to demonstrate that results can be
reproduced when using the original versions, a problem sometimes known
as "code rot."  Researchers will want to know if the results are robust
to the changes. The case studies in [Lapp 2014] provide examples of
these problems.

4. **Barriers to adoption and reuse in existing solutions**

Technological solutions such as workflow software, virtual machines,
continuous integration services, and best practices from software
development would address many of the issues frequently frustrating
reproducibility. However, researchers face siginificant barriers to entry
in learning these tools and approaches which are not part of their typical
curriculum, or lack incentives co-measurate with the effort required
([Joppa 2013], [FitzJohn 2014]).

Existing proposals to address the challenges 1-3 either involve adopting
an relatively new workflow that may be foreign to both the researcher and
their collaborators, as in the case of many workflow software solutions,
or adopt a relatively black-box approach that cannot be easily extended
by other researchers, as in the case of virtual machines.

------------------------------------

<!--
# Docker

At the heart of the computational reproducibility challenge is the 
complexity, diversity, and rapidly changing nature of the computer software
that underpin such work.

-->

# The alternatives

<!-- Current approaches at the Operating Systems level -->

Two dominant paradigms have emerged to address these issues so far:
workflow software (e.g. [Atintas 2004], [Hull 2006])  and
cloud computing on virtual machines (e.g. [Dudley 2013], [Howe 2012]). 
Workflow software provides very elegant technical solutions to the
challenges of communication between diverse software tools, capturing
provenance in graphically driven interfaces, and handling issues from of
versioning dependencies to data access.  Workflow solutions are often 
built by well-funded collaborations between domain scientists and computer
scientists, and can be very successful in the communities within which 
they recieve substatianal adoption. Nonetheless, most workflow systems 
struggle with relatively low total adoption overall ([Gin 2007, Dudley 2013]).

[Dudley 2013] gives several reasons that such comprehensive workflow
systems have not been more successful:

> (i) efforts are not rewarded by the current academic research and
funding environment; (ii) commercial software vendors tend to protect
their markets through proprietary formats and interfaces; (iii)
investigators naturally tend to want to 'own' and control their
research tools; (iv) even the most generalized software will not be
able to meet the needs of every researcher in a field; and finally (v)
the need to derive and publish results as quickly as possible precludes
the often slower standards-based development path.

In short, workflow software expects a new approach to computational research

In contrast, virtual machines (VMs) offer a far more brute-force approach.  If the goal
is to reproduce everything as it ran on the researcher's own computer, well 
then we can just take a snapshot of the entire computer: data, software, 
operating system and all. To make this practical, [Dudley 2013] and [Howe 2012]
propose using virtual machine images that will run on the cloud, such as Amazon's
EC2 system, which is already based upon this kind of virtualization.

Critics of the use of VMs to support reproducibility highlight that the 
approach is too much of a black box and thus ill suited for reproducibility [Watson 2012].
While the approach sidesteps the need to either install or even document the
dependencies, this also makes it more difficult for other researchers to
understand, evaluate, or alter those dependencies.  Moreover, other research
cannot easily build on the virtual machine in a consistent and scalable way. 
If each study provided it's own virtual machine, any pipeline combining the 
tools of multiple studies would quickly become impractical or impossible 
to implement.

## A "DevOpts" approach

The problems highlighted here are not unique to academic software, but
impact software development in general. While the academic research 
literature has frequently focused on the development of workflow software
dedicated to particular domains, or otherwise to the use of virtual machines,
the software development community has recently emphasized a philosophy 
(rather than a particular tool), known as _Development_ and Systems _Operation_, or more 
frequently just "DevOpts." The approach is characterized by scripting,
rather than documenting, a description of the necessary dependencies 
for software to run, usually from the Operating System (OS) on up.

[Clark 2014] describes the DevOpts approach
along with both its relevance to reproducible research and examples of
its use in the academic research context.

Their paper provides both a consise statement of the problem we
have already discussed and the solution provided under this banner:

> In scientific computing the environment was commonly managed via
Makfiles & Unix-y hacks, or alternatively with monolithic software
like Matlab. More recently, centralized package management has provided
curated tools that work well together. But as more and more essential
functionality is built out across a variety of systems and languages,
the value -- and also the difficulty -- of coordinating multiple tools
continues to increase. Whether we are producing research results or web
services, it is becoming increasingly essential to set up new languages,
libraries, databases, and more.

> Documentation for complex software environments is stuck between two
opposing demands. To make things easier on novice users, documentation
must explain details relevant to factors like different operating
systems. Alternatively, to save time writing and updating documentation,
developers like to abstract over such details.

The authors contrast this to the DevOpts approach:

> A DevOps approach to "documenting" an application might consist of
providing brief descriptions of various install paths, along with scripts
or "recipes" that automate setup.

which elegantly addresses both the demand for simplicity of use (one
executes a script instead of manually managing the environmental setup)
and comprehensiveness of implementation. [Clark 2014] are careful to note
that this is not so much a technological shift as a philosophical one:

> The primary shift that's required is not one of new tooling, as most
developers already have the basic tooling they need. Rather, the needed
shift is one of philosophy.

Nevertheless, a growing suite of tools designed explicitly for this
purpose have rapidly replaced the use of general purpose tools (such
as Makefiles, bash scripts) to become synonomous with the "DevOpts"
philosophy.  [Clark 2014] reviews many of these "DevOpts" tools, their
different roles, and their application in reproducible research.

I focus the remainder of this paper on one of the most recent and rapidly
growing among these, called Docker, which offers several promising
developments above and beyond the tools highlighted in [Clark 2014]
in providing a DevOpts approach to reproducible research. My goal
in focusing on this example is not to promote a particular solution, 
but to anchor the discussion of technical solutions to reproducibility
challenges in concrete examples.

<!-- Link on how to install docker -->
<!-- Examples here are not meant as a tutorial to docker, but 
rather to illustrate concepts a reader should be able to follow whether 
or not they have any previous knowledge of docker and its commands -->

# Docker

Docker is an open source (Apache Version 2) software platform written in
the Go programming language.

Docker builds on many long-familiar technologies from operating
systems research: LXC containers, virtualization of the OS, and
a hash-based or git-like verioning and differencing system, among
others. Researchers familiar with LXC may be tempted to view the Docker
system as little more than a creative application and marketing of an
existing technology. However, the Docker approach adds several features
crucial for its use in reproducible research (adapted from [Docker
FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):

- Portable deployment: Docker defines a format for bundling and exchanging
containers such that they run identically across platforms.

- Automatic builds: Docker defines a format (Dockerfiles) that provide
a simple, scripted way to construct containers, following the DevOpts
philosophy.

- Versioning: Docker provides git-like capibilities for tracking changes
to a container, inspecting the history, metadata, or diff between
versions, committing versions, rolling back changes, and incremential
uploads and downloads that send only the differences (like `git push`
or `git pull`).

- Component reuse: Any container can be used as the base for any other.

- Sharing: Docker provides a public registry (Docker Hub) where any
user can make images avaialble to any other. (The registry itself is
also open source so that it can be deployed privately).

Each of these elements have clear implications for reproducible research.

# Using Docker in an R environment

Perhaps the most important feature of a reproducible research tool is
that it be easy to learn and fit relatively seamlessly into exsiting
workflow patterns of domain researchers.  As we have seen, this, more
than any other concern, can explain the relatively low uptake of previously
proposed solutions.  <!-- too strong? -->

Being a new and unfamiliar tool to most domain scientists, Docker is far
from immune to the same critique.  Nevertheless, Docker takes us several 
key steps towards an approach that can be easily adopted in a research 
context.

Primary among these is the ability to run Docker locally on all major
operating systems with minimal prior expertise.  On systems not already
based on the Linux Kernel (such as Mac or Windows clients), this is 
accomplished through the use of a small VirtualBox-based VM running on
the host OS called `boot2docker`.

Local development offers several important advantages in encouraging 
adoption.

-----

In the R environment, the open source RStudio IDE provides another key component
in making this system accessible to most domain scientists. Because the RStudio 
IDE is written completely with standard web languages: javascript, css, and html,
its web-based client, RStudio-Server, looks and feels identical to its popular 
desktop IDE.  RStudio Server provides a way to interact with R on a remote environment
without the latency, X tunnelling, etc typically involved in running R on a remote
server.

For our purposes, it provides users a way to interact with R, the filesystem,
git, text editors, and graphics running on a Docker container.  Users already
familiar with the popular IDE can thus benefit from the reproduciblity and 
portability features provided by running R in a container environment without
having to adapt to a new workflow for managing these things that is usually 
necessary when moving to a cloud environment.

RStudio-Server is not the only way to provide a familiar developer environment
to users working with a development environment running in a container on their
local computer.  Docker supports the sharing of volumes (directories) between
the container and the host OS.  This allows a user to rely on the familiar tools
of the host OS for roles such as text editing, file browsing, or version control,
while still allowing code execution to occur inside the controlled development 
environment of the container. In this example, we launch an interactive R console
in a container that is linked to our current working directory:

```bash
docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
```

(Clearly a similar command could be used just as well with interactive
shells from other languages such as `ipython` or `irb`). The resulting
system behaves almost identically[^2] to running R on the command line
of the host OS.

[^2]: External windows, such graphics windows cannot be opened directly from the container in this setup.  Graphics would have to be saved to desk as raster or vector files and viewed on the host OS, until a better solution can be found. RStudio-server is one way around this.

- Interactive use, scripted use (Dockerfiles)
- Reverting to a previous state
- Integration with local development environment
- Integration with cloud computing
- Transparent & extensible
- serial and linking containers

- Best practices are still taking shape

## Docker remix

The approach of Linux Containers represented by Docker offers a technical
solution to what is frequently seen as the primary weakness of the
standard virtual machine approach to reproducibility - reusing and
remixing elements.  To some extent this is already addressed by the
"DevOpts" approach of Dockerfiles, providing a scripted description
of the environment that can be tweaked and altered, but also includes 
something much more fundamental to Docker.

The challenge to reusing virtual machines can be summarized as "you can't install an image for every pipeline you want..." [Watson 2012]. In contrast, this is exactly how Docker containers are designed to work.
There are at least two ways in which Docker supports this kind
of extensibility.

First, Docker makes it immensely easy to build one container on top of
another.  Rather than copy a Dockerfile and then start adding more lines
to the bottom, we can declare a new Dockerfile is built on an old one
using the `FROM` directive. Here is an example Dockerfile that adds the
R statistical computing environment to a basic Ubuntu Linux distribution.

```
FROM ubuntu:latest
RUN apt-get update && apt-get -y install r-recommended
```

This acts like a software dependency; but unlike other software, a Dockerfile
must have exactly one dependency (one `FROM` line). Note that a particular
version of the dependency can be specified using the `:`.  We can in fact
be much more precise, declaring not only version numbers like `ubuntu:14:04`,
but specific cryptographic hashes that ensure we get the exact same image
every time.

We can now build this Dockerfile and give it name, by running in the working directory:

```bash
docker build -t r-recommended .
```

The key point here is that other researchers can easily build off this 
new image we have just created, extending our work directly, rather than
having to go back to the original image.

However, before others can re-use our image we need a way to share it.

### Docker hub

While Docker images tend to be much smaller than equivalent virtual machines, moving 
around even 100s of gigabytes can be a challenge.  In their work on virtual machines,
[Dudley 2014] recommend only running these tools on cloud servers such as the Amazon
EC2 system. As most researchers still develop software locally and may not have ready
access to these resources, such a requirement adds an additional barrier to reuse.
Fortunately in the case of Docker we have a somewhat more elegant solution through
the use of the Docker Hub [^1].  Docker provides a convenient way to share any image
publicly or privately through the Hub after creating a free account.  Here, we share
a public copy of the image we have just created by using the `docker push` command,
followed by the name of the image:

```bash
docker push username/r-recommended
```

Another user can now build directly on our image, rather than on the `ubuntu:latest`
image we used as a base:

```
FROM username/r-recommended 
RUN apt-get update && apt-get -y install r-cran-teachingdemos
```

[^1]: Similar hubs are in fact available for full virtual machine images, such as the Vagrant hub.

Though we have shown adding only a single piece of software in each step,
clearly this approach can be particularly powerful in building up more complex
environments.

Each one acts as a building block providing just what is necessary to
run one particular service or element, and exposing just what is necessary
to link it together with other blocks.  For instance, we could have one
container running a PostgreSQL database which serves data to another container
running a python environment to analyze the data:

```bash
docker run -d --name db training/postgres
docker run -d -P --link db:db training/webapp python app.py
```

Unlike the much more heavyweight virtual machine approach, containers
are implemented in way such that a single computer can easily run 100s
of such services each in their own container, making it easy to break
computational elements down into logically reusable chunks that come,
batteries included, with everything they need to run reproducibly.
A researcher more familiar with python could connect our MySQL container
to their python container, and so forth.

<!-- too much detail? -->
The implementation of this would nonetheless depend on the individual 
researcher's willingness to break things into multiple containers, as
it is indeed possible to run everything in a monolithic container. 
This is not necessarily bad, as it simplifies the workflow (no need
to execute and link separate containers), and can be refactored into
multiple containers when appropriate. It remains to be seen whether
linking containers through this mechanism

1) Remix.  Titus has an excellent post, "Virtual Machines Considered
Harmful for reproducibility" [1] , .  In contrast,
Docker containers are designed to work exactly like that -- reusable
building blocks you can link together with very little overhead in
disk space or computation.  This more than anything else sets Docker
apart from the standard VM approach.

Docker is not a virtual machine; containers are designed expressly to
be remixable blocks.  You can put an R engine on one container and a
mysql database on another and connect them. Docker philosophy aims at
one function per container to maximize this reuse.  of course it's up
to you to build this way rather than a single monolithic dockerfile,
but the idea of linking containers is a technical concept at the heart
of docker that offers a second and very different way to address the
'remix' problem of VMs.

[Clark 2014]

 Provisioning scripts. Docker images are not 'black boxes'. A
"Dockerfile" is a simple make-like script which installs all the
software necessary to re-create ("provision") the image.  This has
many advantages: (a) The script is much smaller and more portable than
the image. (b) the script can be version managed (c) the script gives
a human readable (instead of binary) description of what software is
installed and how. This also avoids pitfalls of traditional
documentation of dependencies that may be too vague or out-of-sync.
(d) Other users can build on, modify, or extend the script for their
own needs.  All of this is what we call the he "DevOpts" approach to
provisioning, and can be done with AMIs or other virtual machines
using tools like Ansible, Chef, or Puppet coupled with things like
Packer or Vagrant (or clever use of make and shell scripts).

For a much better overview of this "DevOpts" approach in the
reproducible research context and a gentle introduction to these
tools, I highly recommend taking a look at Clark et al [2].

3) You can run the docker container *locally*.  I think this is huge.
In my experience, most researchers do their primary development
locally.  By running RStudio-server on your laptop, it isn't necessary
for me to spin up an EC2 instance (with all the knowledge & potential
cost that requires).  By sharing directories between Docker and the
host OS, a user can still use everything they already know -- their
favorite editor, moving files around with the native OS
finder/browser, using all local configurations, etc, while still
having the code execution occur in the container where the software is
precisely specified and portable.  Whenever you need more power, you
can then deploy the image on Amazon, DigitalOcean, a bigger desktop,
your university HPC cluster, your favorite CI platform, or wherever
else you want your code to run.  (On Mac & Windows, this uses
something called boot2docker, and was not very seamless early on.  It
has gotten much better and continues to improve.)

Related challenges:  RStudio and a consistent environment. 
- Linked volumes 
- linked containers 
- signed containers?

4) Versioned images.  In addition to version managing the Dockerfile,
the images themselves are versioned using a git-like hash system
(check out: docker commit, docker push/pull, docker history, docker
diff, etc).  They have metadata specifying the date, author, parent
image, etc.  We can roll back an image through the layers of history
of its construction, then build off an earlier layer.  This also
allows docker to do all sorts of clever things, like avoiding
downloading redundant software layers from the docker hub.  (If you
pull a bunch of images that all build on ubuntu, you don't get n
copies of ubuntu you have to download and store).  Oh yeah, and
hosting your images on Docker hub is free (no need to pay for an S3
bucket... for now?) and supports automated builds based on your
dockerfiles, which acts as a kind of CI for your environment.
Versioning and diff'ing images is a rather nice reproducibility
feature.

# Further considerations

## Combining virtualization with other reproducible-research tools

<!-- Package managers -->

It is possible, and indeed often desirable, to tackle the problem of dependencies
at a much higher level than that of the operating system itself (e.g. [Ooms 2014]).
For instance, it is often sufficient to address these concerns at the
level of a particular scripting language rather than the level of the
operating system. Most major scripting languages have dedicated tools for
handling dependencies and versioning (e.g. Ruby gems, R packages, Python
eggs, perl modules, javascript node package manager); with some languages
having the fortune or misfortune to have multiple systems. Though systems
differ somewhat in their strategies and their effectiveness, all are
generally vulnerable to changes in their lower-level dependencies (such as
compilers, C libraries, and so forth) that may differ between platforms.
More pointedly, researchers whose work is not easily isolated to a single
such environment but combines tools from different systems (including
combining those tools used for formatting & visual display, such as
LaTeX or pandoc, with scripting or analysis tools such as R).

<!-- Dynamic documents --> 
Dynamic documents are another example of an approach that

approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
the code required to re-generate the results within the manuscript
itself. This process serves as a way to distribute computational
code with publications, to ensure that methodological discriptions,
parameter values, figures, and so forth remain synchronized with the code,
and to provide a way for other researchers to both tweak the code and
'execute' the paper to reproduce the results.  While these features
are very desirable from the perspective of reproducible research,
they inevitably increase the number of software dependencies involved,
including those tools that support combining code with text (e.g. [Xie
2013]) as well as tools merely for the handling the formating and layout
(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
successfully for reasons having nothing to do with the computations
involved in the research, but rather due to software responsible for
implementing these dynamic features.  This problem is largely resolved by
providing a consistent development environmment, making the approaches
discussed here a natural complement to that of dynamic documents.
A similar argument can be made regarding the use of workflow systems,
which can introduce additional softare dependencies even as they seek
to make it easier to work across them.

### Impacting cultural norms?

We noted at the outset that cultural expecations responsible for a
lack of code sharing practices in many fields are a far more extensive
primary barrier to reproducibility than the technical barriers we have
discussed here. Nevertheless, it may be worth considering how solutions
to these technical barriers can influence the cultural landscape as well.
Many researchers may be reluctant to publish code today because they
fear a it will be primarily a one-way street: more technical savvy
researchers then themselves can benefit from their hard work, while
they may not benefit from the work produced by others.  Lowering the
technical barriers to reuse provides immediate practical benefits
that make this exchanged into a more balanced, two-way street. Another
concern is that the difficulty imposed in preparing code to be shared,
such as providing even semi-adequate documentation or support for other
users to be able to install and run it in the first place is too high
[Stodden 2010]. By lowering the barriers these barrier to re-use through
the appropriate infrastructure first, we may also reduce certain cultural
barriers to sharing.

# Acknowledgements

- NSF
- rOpenSci
- Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.

# References

[Garijo 2013]:     http://doi.org/10.1371/journal.pone.0080278
[Lapp 2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
[Gilbert 2012]: 10.1111/j.1365-294X.2012.05754.x
[Collberg 2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
[Brown 2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/

[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
[Ooms 2014]: http://arxiv.org/abs/1406.4806
[Peng 2011]: http://doi.org/10.1126/science.1213847
[Barnes 2010]: http://doi.org/10.1038/467753a
[Ince 2012]: http://doi.org/10.1038/nature10836
[Stodden 2014]: http://doi.org/10.3233/SJI-140818
[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
[Stodden 2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"

[Nature editorial]: http://doi.org/10.1038/483509a
[Nature Special Issue]: http://www.nature.com/nature/focus/reproducibility/

[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."

[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896