### 2014-09-16 manuscript.Rmd
getting there. more writing

---
layout: preprint
title: "An introduction to Docker for reproducible research, with examples from the R environment"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research,
      110 Shaffer Rd,
      Santa Cruz, CA 95050, USA 
abstract: |
  Docker is a promising solution to

bibliography: components/references.bib
csl: components/ecology.csl
documentclass: components/elsarticle

output: 
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

<!--

Suggested paper topics include, but are not limited to:

  * novel ways to manage systems research artifacts, dependencies, workflows,
    and data collection
  * testbeds for repeatable research and/or publishing executable artifacts
  * "rich publications" that connect papers to software and data
  * repeatability in distributed and nondeterministic systems
  * mitigating obstacles to artifact sharing, repeatability, and reproduction
  * repeatability challenges specific to systems
  * measurement bias in systems experiments
  * experiences and outcomes from systems repetition and reproduction studies
  * improving scientific practice in systems research
  * incentives for producing and sharing high-quality research artifacts

Papers are expected to present significant results, insights, and/or directions
for future work, and they are expected to include substantial material that has
not already been published.

SUBMISSION INSTRUCTIONS

Submissions must be no longer than ten (10) pages; shorter papers are allowed
and encouraged.  Submissions must be written in English and follow the standard
formatting guidelines for papers appearing in Operating Systems Review (see
<http://www.sigops.org/osr.html>).  Papers must be submitted in PDF format via
the submission web site.  Submitted papers will be reviewed by the guest editor
and the review committee.  Papers will be evaluated on technical quality,
originality, relevance, and presentation.  Accepted papers will be published in
the January 2015 issue of Operating Systems Review.

-->

Reproducible research has recieved an increasing level of attention
throughout the scientific community (e.g. [Peng 2011], [Nature editorial])
and the public at large (e.g. [The Economist 2014]). All steps of the
scientific process, from data collection and processing, to analyses,
visualizations and conclusions depend ever more on computation and
algorithms, _computational reproducibility_ has recieved particular
attention [Merali 2010].  Though in principle this algorithmic dependence
should make such research easier to reproduce -- computer codes being
both more portable and potentially more precise to exchange and run
than experimental methods -- in practice this has led to an ever larger
and more complex black box that stands between what was actually done
and what is described in the literature.  Crucial scientific processes
such as replicating the results, extending the approach or testing the
conclusions in other contexts, or even merely installing the software
used by the original researchers can become immensely time-consuming if
not impossible.

# A cultural problem

It is worth observing from the outset that the primary barrier to
computational reproducibility in many domain sciences has nothing
to do with the technological approaches discussed here, but stems rather from a
reluctance to publish the code used to generating the results in the
first place ([Barnes 2010]). Despite extensive evidence to the contrary
(e.g. [Ince 2012]), many researchers and journals continue to assume that
summary descriptions or pseudo-code provide a sufficient description of
computational methods used in data gathering, processing, simulation,
visualization, or analysis.  Until such code is available in the first
place, we cannot even begin to encounter the problems that the approaches
discussed here set out to solve. As a result, few domain researchers
may be fully aware of the challenges involved in effectively re-using
published code.

A lack of requirements or incentives no doubt plays a crucial role in
discouraging sharing ([Barnes 2010], [Stodden 2013]). Nevertheless,
it is easy to underestimate the siginficant barriers raised by a lack
of familiar, intuitive, and widely adopted tools for addressing the
challenges of computational reproducibility. Surveys and case studies find
that a lack of time, more than innate opposition to sharing, discourages
researchers from providing code ([Stodden 2010], [FitzJohn 2014]).

# The four technical challenges of computational reproducibility

By restricting ourselves to studies of where code has been made
available, we may sidestep for the moment the cultural challenges
to reproducibility so that we may focus on the technical ones; in
particular, those challenges for which improved tools and techniques
rather than merely norms of behavior can contribute substantially to
improved reproducibility.

Studies focusing on code that has been made available with scientific
publications regularly find the same common issues that pose substantial
barriers to reproducing the original results or building on that code
(e.g. [Lapp 2014], [Garijo 2013], [Gilbert 2012], [Collberg 2014],
[Brown 2014]), which I attempt to summarize as follows.

<!-- Is this a permitted term? -->

1. **"Dependency Hell"**

A recent study by researchers at the University of Arizona found that
less than 50% of software could even be successfully built or installed
[Collberg 2014], and in an ongoing effort by other researchers to
replicate that study [Brown 2014]. Installing or building software
necessary to run the code in question assumes the ability to recreate
the computational environment of the original researchers.

<!-- FIXME more stuff on what this means-->

<!-- Time savings. Personal reproducibility. Collaboration. Teaching. -->

<!-- Figure 1: technical reproducibility stack
1. Operating system
2. Workflow systems 
3. Package managers 
4. Dynamic documents

-->

<!-- Omit this? -->
Differences in numerical evaluation, such as arise in floating point
arithmetic or computers or even ambiguities in standardized programming
languages ("order-of-evaluation" problems) can be responsible for
differing results between or even within the same computational platform
([Ince 2012]). Such issues make it difficult to restrict the true dependencies
of the code to higher level environments such as that of a given scripting
language, indpendent of the underlying OS or even hardware itself.

2.  **Imprecise documentation**

Documentation on how to install and run code associated with published
research is another frequent barrier to replication.  [Lapp 2014] found
this impairs a researcher's ability to install and build the software
necessary, as even small holes in the documentation were found to be
major barriers, particularly for "novices" (Sensu [Garijo 2013] --
where novices may be experts in nearby langauages but unfamiliar with
the package managers and other tools of the language involved). This
same problem is discussed in [Clark 2014].  Imprecise documentation
goes well beyond issues of the software environment itself: incomplete
documentation of parameters involved meant as few as 30% of analyses
(n=34) using the popular software STRUCTURE could be reporoduced in the
study of [Gilbert 2012].

3.  **Code rot**

Software dependencies are not static elements, but recieve regular
updates that may fix bugs, add new features or deprecate old features
(or even entire dependencies themselves).  Any of these changes can
potentially change the results generated by the code.  As some of these
changes may indeed resolve valid bugs or earlier problems with underlying
code, it will often be insufficient to demonstrate that results can be
reproduced when using the original versions, a problem sometimes known
as "code rot."  Researchers will want to know if the results are robust
to the changes. The case studies in [Lapp 2014] provide examples of
these problems.

4. **Barriers to adoption and reuse in existing solutions**

Technological solutions such as workflow software, virtual machines,
continuous integration services, and best practices from software
development would address many of the issues frequently frustrating
reproducibility. However, researchers face siginificant barriers to entry
in learning these tools and approaches which are not part of their typical
curriculum, or lack incentives co-measurate with the effort required
([Joppa 2013], [FitzJohn 2014]).

<!-- Obviously reproducible research barriers go much deeper than this,
code that is not modular, not documented, not functionalized or otherwise
easy to read and follow are all obstacles to resuse.-->

<!-- A wide range of proposals, we focus on those at the operating system level -->
Though a wide variety of approaches exists to work around these challenges,
few operate on a low enough level to provide a general solution.  [Clark 2014]
provide an excellent description of this situation:

> In scientific computing the environment was commonly managed via
Makfiles & Unix-y hacks, or alternatively with monolithic software
like Matlab. More recently, centralized package management has provided
curated tools that work well together. But as more and more essential
functionality is built out across a variety of systems and languages,
the value -- and also the difficulty -- of coordinating multiple tools
continues to increase. Whether we are producing research results or web
services, it is becoming increasingly essential to set up new languages,
libraries, databases, and more.

There are two dominant approaches to this issue of coordinating multiple
tools: Workflows and Virtual Machines (VMs).

# Current approaches

Two dominant paradigms have emerged to address these issues so far:
workflow software (e.g. [Atintas 2004], [Hull 2006])  and
cloud computing on virtual machines (e.g. [Dudley 2013], [Howe 2012]). 
Workflow software provides very elegant technical solutions to the
challenges of communication between diverse software tools, capturing
provenance in graphically driven interfaces, and handling issues from of
versioning dependencies to data access.  Workflow solutions are often 
built by well-funded collaborations between domain scientists and computer
scientists, and can be very successful in the communities within which 
they recieve substatianal adoption. Nonetheless, most workflow systems 
struggle with relatively low total adoption overall ([Gin 2007, Dudley 2013]).

[Dudley 2013] gives several reasons that such comprehensive workflow
systems have not been more successful:

> (i) efforts are not rewarded by the current academic research and
funding environment; (ii) commercial software vendors tend to protect
their markets through proprietary formats and interfaces; (iii)
investigators naturally tend to want to 'own' and control their
research tools; (iv) even the most generalized software will not be
able to meet the needs of every researcher in a field; and finally (v)
the need to derive and publish results as quickly as possible precludes
the often slower standards-based development path.

In short, workflow software expects a new approach to computational research. 
In contrast, virtual machines (VMs) offer a more direct approach. Since
the computer Operating System (OS) already provides the software layer
responsible for coodinating all the different elements running on the computer,
the VM approach captures the OS and everything running on it whole-cloth.
To make this practical, [Dudley 2013] and [Howe 2012]
propose using virtual machine images that will run on the cloud, such as Amazon's
EC2 system, which is already based upon this kind of virtualization.

Critics of the use of VMs to support reproducibility highlight that the 
approach is too much of a black box and thus ill suited for reproducibility [Watson 2012].
While the approach sidesteps the need to either install or even document the
dependencies, this also makes it more difficult for other researchers to
understand, evaluate, or alter those dependencies.  Moreover, other research
cannot easily build on the virtual machine in a consistent and scalable way. 
If each study provided it's own virtual machine, any pipeline combining the 
tools of multiple studies would quickly become impractical or impossible 
to implement.

<!-- Scripted provisioning: a solution to the weaknesses of virtual machines -->
## A "DevOpts" approach

The problems highlighted here are not unique to _academic_ software,
but impact software development in general. While the academic research
literature has frequently focused on the development of workflow software
dedicated to particular domains, or otherwise to the use of virtual
machines, the software development community has recently emphasized a
philosophy (rather than a particular tool), known as _Development_ and
Systems _Operation_, or more frequently just "DevOpts." The approach is
characterized by scripting, rather than documenting, a description of the
necessary dependencies for software to run, usually from the Operating
System (OS) on up.  [Clark 2014] describes the DevOpts approach along
with both its relevance to reproducible research and examples of its
use in the academic research context. They identify the difficulties
we have discussed so far in terms of effective documentation:

> Documentation for complex software environments is stuck between two
opposing demands. To make things easier on novice users, documentation
must explain details relevant to factors like different operating
systems. Alternatively, to save time writing and updating documentation,
developers like to abstract over such details.

The authors contrast this to the DevOpts approach, where dependency 
documentation is _scripted_:

> A DevOps approach to "documenting" an application might consist of
providing brief descriptions of various install paths, along with scripts
or "recipes" that automate setup.

This elegantly addresses both the demand for simplicity of use (one
executes a script instead of manually managing the environmental setup)
and comprehensiveness of implementation. [Clark 2014] are careful to note
that this is not so much a technological shift as a philosophical one:

> The primary shift that's required is not one of new tooling, as most
developers already have the basic tooling they need. Rather, the needed
shift is one of philosophy.

Nevertheless, a growing suite of tools designed explicitly for this
purpose have rapidly replaced the use of general purpose tools (such
as Makefiles, bash scripts) to become synonomous with the "DevOpts"
philosophy.  [Clark 2014] reviews many of these "DevOpts" tools, their
different roles, and their application in reproducible research.

I focus the remainder of this paper on one of the most recent and
rapidly growing among these, called Docker, and the role it can play
in reproducible research. Docker offers several promising features
for reproducibility that go beyond the tools highlighted in [Clark
2014]. Nevertheless, my goal in focusing on this technology is not to
promote a particular solution, but to anchor the discussion of technical
solutions to reproducibility challenges in concrete examples.

<!-- Link on how to install docker -->
<!-- Examples here are not meant as a tutorial to docker, but 
rather to illustrate concepts a reader should be able to follow whether 
or not they have any previous knowledge of docker and its commands -->

# Docker

<!-- Docker is an open source (Apache Version 2) software platform written in
the Go programming language.-->

Docker is an open source project that builds on many long-familiar
technologies from operating systems research: LXC containers,
virtualization of the OS, and a hash-based or git-like verioning
and differencing system, among others.

<!-- Define:
- Dockerfile
- Image 
- Container: When we run a docker image, we get a container.  We can run the same image multiple times, resulting in many identical containers.  We can save the state of a container at any time, which creates a new image.  
- Dockerfile -> build -> image
- Image -> run -> container
-->

I introduce the most relevant concepts from Docker through the 
context of the four challenges for reproducible research I have
discussed above.

### 1. Docker images: resolving 'Dependency Hell'

A Docker based approach works similarly to a virtual machine image in
addressing the dependency problem by providing other researchers with
a binary image in which all the software has already been installed,
configured and tested. (A machine image can also include all data files
necessary for the research, which may simplify the distribution of data.)

A key difference between Docker images and other virutal machines is 
that the Docker images share the Linux kernel with the host machine.
For the end user the primary consequence of this is that any Docker
image must be based on a Linux system with Linux-compatible software, which
includes (R, Python, Matlab, and most other scientific programming needs). [^4]

Sharing the Linux kernel makes Docker much more light-weight and higher
performance than complete virtual machines -- a typical desktop computer
could run no more than a few virutal machines at once but 
would have no trouble running 100s of Docker containers (a container
is simply the term for running instance of an image). This feature
has made Docker particularly attractive to industry and is largely
responsible for the immense popularity of Docker. For our purposes this
is a nice bonus, but the chief value to reproducible research lies
in other aspects.

[^4]: Note that re-distribution of an image in which proprietary software
has been installed will be subject to any relevant licensing agreement.

<!-- 
- image sharing (hub)
- image linking
-->

### 2. Dockerfiles: Resolving imprecise documentation

Though Docker images can be created interactively, this leaves little
transparent record [^3] of what software has been installed and how.
Dockerfiles provide a simple script (similar to a Makefile) that defines
exactly how to build up the image, consistent with the DevOpts approach
we mentioned previously.

With a syntax that is simpler than other provisioning tools (e.g. Chef,
Puppet, Ansible) or Continuous Integration (CI) platforms (e.g. Travis CI,
Shippable CI); users need little more than a basic familarity with shell
scripts and a Linux distribution software environment (e.g. Debian-based
`apt-get`) to get started writing Dockerfiles.

This approach has many advantages:

- While machine images can be very large (many gigabytes), a Dockerfile
is just a small plain text file that can be easily stored and shared.

- Small plain text files are ideally suited for use with a version
management system such as `subversion` or `git`, which can track any
changes made to the `Dockerfile`

- the `Dockerfile` provides a human readable summary of the necessary
software dependencies, environmental variables and so forth needed to
execute the code. There is little possiblity of the kind of holes or 
imprecision in such a script that so frequently cause difficulty in
manually implemented documentation of dependencies. This approach
also avoids the burden of having to tediously document dependencies 
at the end of a project, since they are instead documented as they 
are installed by writing the `Dockerfile`.

- Unlike a `Makefile` or other script, the `Dockerfile` includes all
software dependencies down to the level of the OS, and is built by the
Docker `build` tool, making it very unlikely that the resulting build
will differ when being built on different machines. This is not to say
that all builds of a Dockerfile are bitwise identical.  In particular,
builds executed later will install more recent versions of the same
software, if available, unless the package managers used are explicitly
configured otherwise. We address this issue in the next section.

- It is possible to add checks and tests following the commands for
installing the software environment, which will verify that the setup
has been successful.  This can be important in addressing the issue of
code-rot which we discuss next.

- It is straight-straight forward for other users to extend or customize 
the resulting image by editing the script directly.

[^3]: The situation is in fact slightly better than the virtual machine
approach because these changes are versioned. Docker provides tools
to inspect differences (diffs) between the images, and we can also roll
back changes to earlier versions.

### 3. Tackling code-rot with image versions

As I have discussed above, changes to the dependencies, whether they are
the result of security fixes, new features, or deprecation of old software,
can break otherwise functioning code. These challenges can be significantly
reduced because Docker defines the software environment to a particular operating system 
and suite of libraries, such as the Ubuntu or Debian distribution.  Such 
distributions use a staged release model with `stable`, `testing` and `unstable` 
phases subjected to extensive testing to catch such potential problems [Ooms 2013],
while also providing regular security updates to software within each stage.
Nonetheless, this cannot completely avoid the challenge of code-rot,
particularly when it is necessary to install software that is not (yet)
available for a given distribution.

To address this concern, one must archive a binary copy of the image
used at the time the research was first performed.  Docker provides a
simple utility to save an image as a portable `tarball` file that can
be read in by any other Docker installation, providing a robust way to
run the exact versions of all software involved.  By testing both the
`tarball` archive and the image generated by the latest Dockerfile,
Docker provides a simple way to confirm whether or not code rot has
effected the function of a particular piece of code.

<!-- FIXME move this later --> 
To simplify this process, Docker also supports *Automated Builds*
through the Docker Hub.  This acts as a kind of Continous Integration
(CI) service that verifies the image builds correctly whenever the
Dockerfile is updated, particularly if the Dockerfile includes checks
for the environment. The Hub also provides a convenient distribution
service, freely storing the prebuilt images, along with their metadata,
for download and reuse by others. The Docker Hub is a free service and
an open source software product so that users can run their own private
versions of the Hub on their own servers, for instance, if security of
the data or the longevity of the public platform is a concern.

### 4. Barriers to adoption and re-use

A technical solution, no matter how elegant, will be of little practical 
use for reproducible research unless it is both easy to use and adapt to
the existing workflow patterns of practicing domain researchers.

Though most of the concerns we have discussed so far can be addressed
through well-designed workflow software or the use of a DevOpts approach
to provisioning virtual machines by scripts, neither approach has seen
widespread adoption by domain researchers, who work primarily in a
local rather than cloud-based environment using develoment tools native
to their personal operating system. To gain more widespread adoption,
reproducible research technologies must make it easier, not harder,
for a researcher to perform the tasks they are already doing (before 
considering any additional added benefits).

These issues are reflected both during the original research or development
phase and in any subsequent reuse. Another researcher may be less likely
to build on existing work if it can only be done by using a particular
workflow system or monolothic software platform with which they are
unfamiliar.  Likewise, a user is more likey to make their own computational
environment avialable for reuse if it does not involve a significant
added effort in packaging and documenting [Stodden 2010].

Though Docker is not immune to these challenges, it offers a interesting 
example of a way forward in addressing these fundamental cocerns.

- *Local environment*
- *Modular reuse*

<!---
- The ability to work in a familiar, local environment (during one's own research or when re-using elements from another researcher)
- An Docker container approach to both development and deployment
- Image re-use & Image linking 
-->

# Using Docker as a local development environment

Perhaps the most important feature of a reproducible research tool is
that it be easy to learn and fit relatively seamlessly into exsiting
workflow patterns of domain researchers.  As we have seen, this, more than
any other concern, can explain the relatively low uptake of previously
proposed solutions.  <!-- too strong? -->

Being a new and unfamiliar tool to most domain scientists, Docker is
far from immune to the same critique.  Nevertheless, Docker takes us
several key steps towards an approach that can be easily adopted in a
research context.

Primary among these is the ability to run Docker locally on all major
operating systems with minimal prior expertise.  On systems not already
based on the Linux Kernel (such as Mac or Windows clients), this is
accomplished through the use of a small VirtualBox-based VM running on
the host OS called `boot2docker`.

Local development offers several important advantages in encouraging
adoption.

- Shared volumes
- Host OS tools

-----

In the R environment, the open source RStudio IDE provides
another key component in making this system accessible to most
domain scientists. Because the RStudio IDE is written completely with
standard web languages (javascript, css, and html), its web-based 
RStudio Server looks and feels identical to its popular desktop IDE.
RStudio Server provides a way to interact with R on a remote environment
without the latency, X tunnelling, etc typically involved in running R
on a remote server.

For our purposes, it provides users a way to interact with R,
the filesystem, git, text editors, and graphics running on a Docker
container.  Users already familiar with the popular IDE can thus benefit
from the reproduciblity and portability features provided by running R
in a container environment without having to adapt to a new workflow
for managing these things that is usually necessary when moving to a
cloud environment.

RStudio-Server is not the only way to provide a familiar developer
environment to users working with a development environment running in
a container on their local computer.  Docker supports the sharing of
volumes (directories) between the container and the host OS.  This allows
a user to rely on the familiar tools of the host OS for roles such as
text editing, file browsing, or version control, while still allowing
code execution to occur inside the controlled development environment
of the container. In this example, we launch an interactive R console
in a container that is linked to our current working directory:

```bash
docker run -v $(pwd):/ -it eddelbuettel/docker-ubuntu-r /usr/bin/R
```

(Clearly a similar command could be used just as well with interactive
shells from other languages such as `ipython` or `irb`). The resulting
system behaves almost identically[^2] to running R on the command line
of the host OS.

[^2]: External windows, such graphics windows cannot be opened directly
from the container in this setup.  Graphics would have to be saved to
desk as raster or vector files and viewed on the host OS, until a better
solution can be found. RStudio-server is one way around this.

## Docker remix: Building re-usable modules

The approach of Linux Containers represented by Docker offers a technical
solution to what is frequently seen as the primary weakness of the
standard virtual machine approach to reproducibility - reusing and
remixing elements.  To some extent this is already addressed by the
"DevOpts" approach of Dockerfiles, providing a scripted description
of the environment that can be tweaked and altered, but also includes
something much more fundamental to Docker.

The challenge to reusing virtual machines can be summarized as "you
can't install an image for every pipeline you want..." [Watson 2012]. In
contrast, this is exactly how Docker containers are designed to work.
There are at least two ways in which Docker supports this kind of
extensibility.

First, Docker makes it immensely easy to build one container on top of
another.  Rather than copy a Dockerfile and then start adding more lines
to the bottom, we can declare a new Dockerfile is built on an old one
using the `FROM` directive. Here is an example Dockerfile that adds the
R statistical computing environment to a basic Ubuntu Linux distribution.

```
FROM ubuntu:latest
RUN apt-get update && apt-get -y install r-recommended
```

This acts like a software dependency; but unlike other software, a
Dockerfile must have exactly one dependency (one `FROM` line). Note that
a particular version of the dependency can be specified using the `:`.
We can in fact be much more precise, declaring not only version numbers
like `ubuntu:14:04`, but specific cryptographic hashes that ensure we
get the exact same image every time.

We can now build this Dockerfile and give it name, by running in the
working directory:

```bash
docker build -t r-recommended .
```

The key point here is that other researchers can easily build off this 
new image we have just created, extending our work directly, rather than
having to go back to the original image.

However, before others can re-use our image we need a way to share it.

### Docker hub

While Docker images tend to be much smaller than equivalent virtual
machines, moving around even 100s of gigabytes can be a challenge.
In their work on virtual machines, [Dudley 2014] recommend only
running these tools on cloud servers such as the Amazon EC2 system. As
most researchers still develop software locally and may not have ready
access to these resources, such a requirement adds an additional barrier
to reuse.  Fortunately in the case of Docker we have a somewhat more
elegant solution through the use of the Docker Hub [^1].  Docker provides
a convenient way to share any image publicly or privately through the
Hub after creating a free account.  Here, we share a public copy of the
image we have just created by using the `docker push` command, followed
by the name of the image:

```bash
docker push username/r-recommended
```

Another user can now build directly on our image, rather than on the `ubuntu:latest`
image we used as a base:

```
FROM username/r-recommended 
RUN apt-get update && apt-get -y install r-cran-teachingdemos
```

[^1]: Similar hubs are in fact available for full virtual machine images, such as the Vagrant hub.

Though we have shown adding only a single piece of software in each step,
clearly this approach can be particularly powerful in building up more complex
environments.

Each one acts as a building block providing just what is necessary to
run one particular service or element, and exposing just what is necessary
to link it together with other blocks.  For instance, we could have one
container running a PostgreSQL database which serves data to another container
running a python environment to analyze the data:

```bash
docker run -d --name db training/postgres
docker run -d -P --link db:db training/webapp python app.py
```

Unlike the much more heavyweight virtual machine approach, containers
are implemented in way such that a single computer can easily run 100s
of such services each in their own container, making it easy to break
computational elements down into logically reusable chunks that come,
batteries included, with everything they need to run reproducibly.
A researcher could connect a container providing a computational 
environment for a different language to this same MySQL container, 
and so forth.

<!-- too much detail? -->
The implementation of this would nonetheless depend on the individual 
researcher's willingness to break things into multiple containers, as
it is indeed possible to run everything in a monolithic container. 
This is not necessarily bad, as it simplifies the workflow (no need
to execute and link separate containers), and can be refactored into
multiple containers when appropriate. It remains to be seen whether
linking containers through this mechanism

----------------

- Interactive use, scripted use (Dockerfiles)
- Reverting to a previous state
- Integration with local development environment
- Integration with cloud computing
- Transparent & extensible
- serial and linking containers
- Best practices are still taking shape

----------

##  Best practices

- Using dockerfiles instead of containers
- Using appropriate, common base images
- Archiving tarball snapshots
- Adding tests/checks to the Dockerfile
- Sharing 
- Leveraging package managers appropriately to handle software dependencies

Upcoming features:

- signed containers?

-------------

3) You can run the docker container *locally*.  I think this is huge.
In my experience, most researchers do their primary development
locally.  By running RStudio-server on your laptop, it isn't necessary
for me to spin up an EC2 instance (with all the knowledge & potential
cost that requires).  By sharing directories between Docker and the
host OS, a user can still use everything they already know -- their
favorite editor, moving files around with the native OS
finder/browser, using all local configurations, etc, while still
having the code execution occur in the container where the software is
precisely specified and portable.  Whenever you need more power, you
can then deploy the image on Amazon, DigitalOcean, a bigger desktop,
your university HPC cluster, your favorite CI platform, or wherever
else you want your code to run.  (On Mac & Windows, this uses
something called boot2docker, and was not very seamless early on.  It
has gotten much better and continues to improve.)

4) Versioned images.  In addition to version managing the Dockerfile,
the images themselves are versioned using a git-like hash system
(check out: docker commit, docker push/pull, docker history, docker
diff, etc).  They have metadata specifying the date, author, parent
image, etc.  We can roll back an image through the layers of history
of its construction, then build off an earlier layer.  This also
allows docker to do all sorts of clever things, like avoiding
downloading redundant software layers from the docker hub.  (If you
pull a bunch of images that all build on ubuntu, you don't get n
copies of ubuntu you have to download and store).  Oh yeah, and
hosting your images on Docker hub is free (no need to pay for an S3
bucket... for now?) and supports automated builds based on your
dockerfiles, which acts as a kind of CI for your environment.
Versioning and diff'ing images is a rather nice reproducibility
feature.

--------------------------------

<!-- Covered elsewhere in a more integrated fashion -->
Researchers familiar with LXC may be tempted to view the Docker
system as little more than a creative application and marketing of an
existing technology. However, the Docker approach adds several features
crucial for its use in reproducible research (adapted from [Docker
FAQ](https://docs.docker.com/faq/#what-does-docker-add-to-just-plain-lxc)):

- *Portable deployment*: Docker defines a format for bundling and
exchanging containers such that they run identically across platforms.

- *Automatic builds*: Docker defines a format (Dockerfiles) that provide
a simple, scripted way to construct containers, following the DevOpts
philosophy.

- *Versioning*: Docker provides git-like capibilities for tracking
changes to a container, inspecting the history, metadata, or diff between
versions, committing versions, rolling back changes, and incremential
uploads and downloads that send only the differences (like `git push`
or `git pull`).

- *Component reuse*: Any container can be used as the base for any other.

- *Sharing*: Docker provides a public registry (Docker Hub) where any
user can make images avaialble to any other. (The registry itself is
also open source so that it can be deployed privately).

Each of these elements have clear implications for reproducible research.

# Further considerations

## Combining virtualization with other reproducible-research tools

<!-- Package managers -->

It is possible, and indeed often desirable, to tackle the problem of
dependencies at a much higher level than that of the operating system
itself (e.g. [Ooms 2014]).  For instance, it is often sufficient to
address these concerns at the level of a particular scripting language
rather than the level of the operating system. Most major scripting
languages have dedicated tools for handling dependencies and versioning
(e.g. Ruby gems, R packages, Python eggs, perl modules, javascript node
package manager); with some languages having the fortune or misfortune to
have multiple systems. Though systems differ somewhat in their strategies
and their effectiveness, all are generally vulnerable to changes in their
lower-level dependencies (such as compilers, C libraries, and so forth)
that may differ between platforms.  More pointedly, researchers whose
work is not easily isolated to a single such environment but combines
tools from different systems (including combining those tools used for
formatting & visual display, such as LaTeX or pandoc, with scripting or
analysis tools such as R).

<!-- Dynamic documents --> Dynamic documents are another example of an
approach that

approach ([Leisch 2002], [Peng 2011], [Xie 2013]), which seek to embed
the code required to re-generate the results within the manuscript
itself. This process serves as a way to distribute computational
code with publications, to ensure that methodological discriptions,
parameter values, figures, and so forth remain synchronized with the code,
and to provide a way for other researchers to both tweak the code and
'execute' the paper to reproduce the results.  While these features
are very desirable from the perspective of reproducible research,
they inevitably increase the number of software dependencies involved,
including those tools that support combining code with text (e.g. [Xie
2013]) as well as tools merely for the handling the formating and layout
(e.g. LaTeX, pandoc).  Consequently, such a paper can fail to execute
successfully for reasons having nothing to do with the computations
involved in the research, but rather due to software responsible for
implementing these dynamic features.  This problem is largely resolved by
providing a consistent development environmment, making the approaches
discussed here a natural complement to that of dynamic documents.
A similar argument can be made regarding the use of workflow systems,
which can introduce additional softare dependencies even as they seek
to make it easier to work across them.

### Impacting cultural norms?

We noted at the outset that cultural expecations responsible for a
lack of code sharing practices in many fields are a far more extensive
primary barrier to reproducibility than the technical barriers we have
discussed here. Nevertheless, it may be worth considering how solutions
to these technical barriers can influence the cultural landscape as well.
Many researchers may be reluctant to publish code today because they
fear a it will be primarily a one-way street: more technical savvy
researchers then themselves can benefit from their hard work, while
they may not benefit from the work produced by others.  Lowering the
technical barriers to reuse provides immediate practical benefits
that make this exchanged into a more balanced, two-way street. Another
concern is that the difficulty imposed in preparing code to be shared,
such as providing even semi-adequate documentation or support for other
users to be able to install and run it in the first place is too high
[Stodden 2010]. By lowering the barriers these barrier to re-use through
the appropriate infrastructure first, we may also reduce certain cultural
barriers to sharing.

# Acknowledgements

- NSF
- rOpenSci
- Dirk Eddlebuettel, Rich FitzJohn, Yihui Xie, Titus Brown, John Stanton-Geddes and others for helpful discussions about reproducibility, virtualization, and Docker that have helped shape this manuscript.

# References

[Garijo 2013]:     http://doi.org/10.1371/journal.pone.0080278
[Lapp 2014]: https://storify.com/hlapp/reproducibility-repeatability-bigthink
[Gilbert 2012]: 10.1111/j.1365-294X.2012.05754.x
[Collberg 2014]: http://reproducibility.cs.arizona.edu/v1/tr.pdf
[Brown 2014]: http://cs.brown.edu/~sk/Memos/Examining-Reproducibility/

[Vandewalle 2009]: http://doi.org/10.1109/MSP.2009.932122
[Merali 2010]: http://doi.org/10.1038/467775a "Computational Science ... error: why scientific programming does not compute"
[Gil 2007]: http://doi.org/10.1109/MC.2007.421 "Examining the Challenges of Scientific Workflows"
[Hull 2006]: http://doi.org/10.1093/nar/gkl320 "Taverna: a tool for building and running workflows of services"
[Altintas 2004]: http://doi.org/10.1109/SSDM.2004.1311241 "Kepler: an extensible system for design and execution of scientific workflows"
[Howe 2012]: http://doi.org/10.1109/MCSE.2012.62
[Dudley 2013]: http://doi.org/10.1038/nbt1110-1181 
[Clark 2014]: https://berkeley.app.box.com/s/w424gdjot3tgksidyyfl
[Ooms 2013]: http://arxiv.org/abs/1303.2140v2
[Ooms 2014]: http://arxiv.org/abs/1406.4806
[Peng 2011]: http://doi.org/10.1126/science.1213847
[Barnes 2010]: http://doi.org/10.1038/467753a
[Ince 2012]: http://doi.org/10.1038/nature10836
[Stodden 2014]: http://doi.org/10.3233/SJI-140818
[Stodden 2010]: http://doi.org/10.2139/ssrn.1550193
[Stodden 2013]: http://www.davidhbailey.com/dhbpapers/icerm-report.pdf "ICERM Working group report: Setting the Default to Reproducible: Reproducibility in Computational and Experimental Mathematics"

[Nature editorial]: http://doi.org/10.1038/483509a
[Nature Special Issue]: http://www.nature.com/nature/focus/reproducibility/

[van der Meulen 2008]: http://doi.org/10.1109/TSE.2008.70 "The effectiveness of software diversity in a large population of programs."

[Watson 2012]: https://twitter.com/BioMickWatson/status/265037994526928896